---
title: "Color Score Analysis for Dani"
author: "EL Strand"
output:
  github_document: default
  pdf_document:
    keep_tex: yes
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

## Loading in required libraries 

```{r}
library(plyr)
library("vegan")
library(dplyr)
library(tidyverse)
library(readxl)
library(writexl)
library(Rmisc)
```

## Read in data

```{r}
tropical_data <- read_excel("Tropical_heatwave_color_score_data.xlsx", sheet = "TROPICAL",
                            col_types = c("text", "text", "text",
                                          "numeric", "numeric", "numeric",
                                          "numeric", "numeric", "numeric",
                                          "numeric", "text", "text"))

temperate_data <- read_excel("Tropical_heatwave_color_score_data.xlsx", sheet = "TEMPERATE",
                            col_types = c("text", "text", "text",
                                          "numeric", "numeric", "numeric",
                                          "numeric", "numeric", "numeric",
                                          "numeric", "text"))


```

## Create sample ID column 

```{r}
tropical_data <- tropical_data %>% 
  unite(Sample, c("Timepoint", "Tank_Number", "Fragment_Number"), sep = " ", remove = FALSE) 

temperate_data <- temperate_data %>% 
  unite(Sample, c("Timepoint", "Tank_Number", "Fragment_Number"), sep = " ", remove = FALSE) 
```


## Normalizing to color standards 

Tropical set 

```{r}
tropical_data$Red.Norm.Coral <- tropical_data$Red_coral/tropical_data$Red_standard
tropical_data$Green.Norm.Coral <- tropical_data$Green_coral/tropical_data$Green_standard
tropical_data$Blue.Norm.Coral <- tropical_data$Blue_coral/tropical_data$Blue_standard

head(tropical_data)
```

Temperate set

```{r}
temperate_data$Red.Norm.Coral <- temperate_data$Red_coral/temperate_data$Red_standard
temperate_data$Green.Norm.Coral <- temperate_data$Green_coral/temperate_data$Green_standard
temperate_data$Blue.Norm.Coral <- temperate_data$Blue_coral/temperate_data$Blue_standard

head(temperate_data)
```

## Principal Components Analysis to find 'Color Score' value 

### Create data matrix and name columns 

Tropical set 

```{r}
tropical_matrix <- as.matrix(cbind(tropical_data$Red.Norm.Coral,
                                   tropical_data$Green.Norm.Coral,
                                   tropical_data$Blue.Norm.Coral)) #create matrix

rownames(tropical_matrix) <- tropical_data$Sample #name columns in dataframe
```

Temperate set 

```{r}
temperate_matrix <- as.matrix(cbind(temperate_data$Red.Norm.Coral,
                                   temperate_data$Green.Norm.Coral,
                                   temperate_data$Blue.Norm.Coral)) #create matrix

rownames(temperate_matrix) <- temperate_data$Sample #name columns in dataframe
```

### Calculate distance matrix of color scores

```{r}
tropical_dist <- vegdist(tropical_matrix, method="euclidean") 
temperate_dist <- vegdist(temperate_matrix, method="euclidean") 
```

### Calculate PCA 

```{r}
trop_PCA <- princomp(tropical_dist) #run principal components Analysis
summary(trop_PCA) # view variance explained by PCs

temp_PCA <- princomp(temperate_dist)
summary(temp_PCA)
```

### Extract PC1 as color score value

Tropical set

```{r}
Tropical_colorscore <- as.data.frame(-trop_PCA$scores[,1]) #extract PC1
Tropical_colorscore$Sample <- rownames(tropical_matrix)

Tropical_colorscore <- Tropical_colorscore %>% dplyr::rename(., Color_score = `-trop_PCA$scores[, 1]`)
```

Temperate set 

```{r}
Temperate_colorscore <- as.data.frame(-temp_PCA$scores[,1]) #extract PC1
Temperate_colorscore$Sample <- rownames(temperate_matrix)

Temperate_colorscore <- Temperate_colorscore %>% dplyr::rename(., Color_score = `-temp_PCA$scores[, 1]`)
```

## Create final data frame 

```{r}
tropical_data_final <- left_join(tropical_data, Tropical_colorscore, by = "Sample") 

temperate_data_final <- left_join(temperate_data, Temperate_colorscore, by = "Sample")
```

## Export that data frame 

```{r}
tropical_data_final %>% write_xlsx("Tropical_colorscore.xlsx")
temperate_data_final %>% write_xlsx("Temperate_colorscore.xlsx")
```

# Analysis 

## Import data frame from the above so don't have to run code again 

```{r}
tropical_df <- read_excel("Tropical_colorscore.xlsx") %>% 
  select(1:4,17) %>%
  mutate(Treatment = case_when(Tank_Number == "1" ~ "Ambient Temperature", 
                               Tank_Number == "2" ~ "High Temperature",
                               Tank_Number == "3" ~ "Ambient Temperature",
                               Tank_Number == "4" ~ "High Temperature")) %>%
  mutate(Date = case_when(Timepoint == "20220329" ~ "Day 1", 
                          Timepoint == "20220331" ~ "Day 3",
                          Timepoint == "20220402" ~ "Day 6",
                          Timepoint == "20220404" ~ "Day 8",
                          Timepoint == "20220406" ~ "Day 10",
                          Timepoint == "20220408" ~ "Day 12",
                          Timepoint == "20220411" ~ "Day 15",
                          Timepoint == "20220414" ~ "Day 18",
                          Timepoint == "20220417" ~ "Day 21",
                          Timepoint == "20220419" ~ "Day 23"))

tropical_df$Date <- factor(tropical_df$Date, 
                           levels = c("Day 1", "Day 3", "Day 6", "Day 8",
                                      "Day 10", "Day 12", "Day 15", "Day 18",
                                      "Day 21", "Day 23"))

tropical_df %>% write_xlsx("Tropical_colorscore_meta.xlsx")

temperate_df <- read_excel("Temperate_colorscore.xlsx") %>% 
  select(2:4,16) %>% 
  mutate(Treatment = case_when(Tank_Number == "1" ~ "High Temperature", 
                               Tank_Number == "2" ~ "Ambient Temperature",
                               Tank_Number == "3" ~ "High Temperature",
                               Tank_Number == "4" ~ "Ambient Temperature")) %>%
  filter(Color_score > -100) %>% ## this threshold was decided after looking at the histograms below
  mutate(Date = case_when(Timepoint == "20220914" ~ "Baseline", 
                          Timepoint == "20220921" ~ "Day 2",
                          Timepoint == "20220923" ~ "Day 4",
                          Timepoint == "20220927" ~ "Day 8",
                          Timepoint == "20220929" ~ "Day 10",
                          Timepoint == "20220930" ~ "Day 11",
                          Timepoint == "20221004" ~ "Day 15",
                          Timepoint == "20221006" ~ "Day 17"))

temperate_df$Date <- factor(temperate_df$Date, 
                           levels = c("Baseline", "Day 2", "Day 4", "Day 8",
                                      "Day 10", "Day 11", "Day 15", "Day 17"))

temperate_df %>% write_xlsx("Temperate_colorscore_meta.xlsx")
```

```{r}
hist(tropical_df$Color_score)
hist(temperate_df$Color_score)
```

```{r}
tropical_means <- summarySE(tropical_df, measurevar = c("Color_score"), groupvars = c("Date", "Treatment"))
temperate_means <- summarySE(temperate_df, measurevar = c("Color_score"), groupvars = c("Date", "Treatment"))
```

```{r}
tropical_means %>%
  ggplot(., aes(x=Date, y=Color_score, group=Treatment, color=Treatment)) +
  geom_point() + geom_line() + 
  scale_color_manual(values = c("blue", "red")) +
  geom_errorbar(aes(ymin=Color_score-se, ymax=Color_score+se), width=.2) +
  theme_bw() + ylab("Color Score") + 
  xlab("")

temperate_means %>%
  ggplot(., aes(x=Date, y=Color_score, group=Treatment, color=Treatment)) +
  geom_point() + geom_line() + 
  scale_color_manual(values = c("blue", "red")) +
  geom_errorbar(aes(ymin=Color_score-se, ymax=Color_score+se), width=.2) +
  theme_bw() + ylab("Color Score") + 
  xlab("")
```



