---
title: "BiocSwirl_courses"
author: "EL Strand"
date: "1/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clears workspace. 
```{r}
rm(list=ls())
```

Loading libraries. I had trouble downloading devtools so I updated 'processx' package, installed 'usethis', and re-ran the install devtools function. Seems to have fixed this.  
1/20/21: devtools wouldn't run without ellipsis. Installed ellipsis package, 'yes' to sources that need compilation.
```{r}
install.packages("devtools")
library(devtools)

devtools::install_github("biocswirl-dev-team/BiocSwirl") # selected '1' to update all; 'Yes' to install that need compilation; 'Yes' to install from C/C++/Fortran: 'brio'. This install took awhile. 
library(BiocSwirl)
```

```{r, eval = FALSE}
# list all available BiocSwirl courses
list_courses()
# install selected course. 
# copy name from output of list_courses()
load_course(course ='RNAseq')
```

To start the course environment, excute the command below. No parameters are necessary. The command will start an interactive environment that will allow you to navigate to your course.
```{r, eval = FALSE}
start_course()
```

## RNASeq Course

Installing packages needed for analysis. 
```{r}
install.packages("pheatmap")
install.packages("BiocManager")
BiocManager::install("DESeq2")
```


